US CES Analysis for Victor Court

### 2018-07-11 Articles/Court_Analysis_2018/FittingForCourt2018.Rmd
Report for Victor.

---
title: "Analysis for Victor Court"
author: "Matthew Kuperus Heun"
date: "10 July 2018"
output: pdf_document
---

@@

Victor has asked me to use my CES fitting functions on his US data
as a point of comparison.

@@

```{r}
USData <- USData_raw %>% 
  select(-K.L...US.2010.per.hour., 
         -Capital...US.2010., 
         -Labor..millions.of.hours., 
         -Primary.exergy..TJ.,
         -GDP....US.2010.) %>% 
  rename(
    k = K..1960...1.,
    l = L..1960...1.,
    e = E..1960...1.,
    y = Y..1960...1.
  ) %>% 
  mutate(
    Country = "US"
  ) %>% 
  select(Country, Year, y, k, l, e)
head(USData, n = 3)
```

Now review the data.

@@

We'll fit with the $(kl)e$ nesting.

```{r}
kleModel <- cesModel(y ~ k + l + e, data = USData)
kleModel
nckle <- naturalCoef(kleModel)
nckle
```

Note that $\delta = 1$, so this is a boundary model.
We also have $\rho_1 = -1$ (therefore, $\sigma_1 = \infty$). 
We can query which boundary model we have: 
```{r}
kleModel$bname
```
So this is boundary model 7 of Table 2 in the ``Cautionary Tales'' paper.

With $\theta = \mathrm{e}^{\mathrm{logscale}} = \mathrm{e}^
{`r naturalCoef(kleModel)[["logscale"]]`} =
`r naturalCoef(kleModel)[["scale"]]`$,
these results mean that boundary model 7

$$y = \theta A \left[ \delta_1 x_1 + (1 - \delta_1) x_2 \right]$$
is filled with coefficients like this:

$$y = `r exp(coef(kleModel)[["logscale"]])` 
      \mathrm{e}^{-`r coef(kleModel)[["lambda"]]` (t-1960)} 
      \left[ `r coef(kleModel)[["delta_1"]]` \, k 
          + `r 1 - coef(kleModel)[["delta_1"]]` \, l \right] \; .$$

@@

We can compare the fitted model against the historical data
as shown in the following graph.

```{r}
USData %>% 
  mutate(
    `y predicted` = yhat(kleModel) %>% as.numeric() # avoids an attributes warning
  ) %>% 
  select(Year, y, `y predicted`) %>% 
  gather(key = var, value = value, y, `y predicted`) %>% 
  ggplot(mapping = aes(x = Year, y = value, linetype = var)) +
  geom_line() + 
  geom_hline(yintercept = 1, colour = "gray50", size = 0.1) +
  labs(x = NULL, y = NULL, linetype = NULL) +
  xy_theme()
```

@@

1. The fit is rather good.
2. It is likely that the fitted model excludes energy ($e$), 
   because the energy dip in 1980 is too large. 
   Instead, by picking up the smaller dip in the $l$ data series, the fit is better.
3. I find it interesting, but don't quite know what to make of it, that
   the coefficients on $k$ and $l$ are approximately the neoclassical numbers 
   (0.7 and 0.3, respectively).

>>>>

---
title: "US CES Analysis for Victor Court"
author: "Matthew Kuperus Heun, Calvin College"
date: "11 July 2018"
output: pdf_document
---

@@

Victor Court has asked me to fit US energy and economic data
using the CES aggregate production function. 
The energy and economic data have been supplied by Victor.
I will use the fitting code that was created for the ``Cautionary Tales'' paper.

@@

```{r}
USData <- USData_raw %>% 
  select(-K.L...US.2010.per.hour., 
         -Capital...US.2010., 
         -Labor..millions.of.hours., 
         -Primary.exergy..TJ.,
         -GDP....US.2010.) %>% 
  rename(
    k = K..1960...1.,
    l = L..1960...1.,
    e = E..1960...1.,
    y = Y..1960...1.
  ) %>% 
  mutate(
    Country = "US",
    time = Year - min(Year)
  ) %>% 
  select(Country, Year, time, y, k, l, e) # Re-order the columns
head(USData, n = 3)
```

Now review the data in graphical form.

@@

We'll fit with the $(kl)e$ nesting.
The warning occurs in one of the many trials and can be ignored.

```{r}
kleModel <- cesModel(y ~ k + l + e + time, data = USData)
kleModel
```
Note that $\rho_1 = -1$ (therefore, $\sigma_1 = \infty$). 
Therefore, this is a boundary model.

We can query which boundary model we have obtained: 
```{r}
kleModel$bname
```
So this is boundary model 20 of Table 2 in the ``Cautionary Tales'' paper.

We can see the "natural" coefficients for the model,
which shows that the "scale" parameter ($\theta$) is close to 1, as expected.
```{r}
naturalCoef(kleModel)
```

We can also try other nestings of the CES model, 
which provide three degenerate cases: $(kl)$, $(le)$, and $(ek)$ nestings.
In all of these, `delta` is 1 and `sigma` and `rho` are `NA`, 
because the third factor of production is missing.

```{r}
klModel <- cesModel(y ~ k + l + time, data = USData)
naturalCoef(klModel)
leModel <- cesModel(y ~ l + e + time, data = USData)
naturalCoef(leModel)
ekModel <- cesModel(y ~ e + k + time, data = USData)
naturalCoef(ekModel)
```

@@

We can compare the fitted models against the historical data
as shown in the following graph.
```yhat``` is a function that extracts the fitted values.

```{r}
USData %>% 
  mutate(
    `y fitted kle` = yhat(kleModel) %>% as.numeric(), # avoids an attributes warning
    `y fitted kl` = yhat(klModel),
    `y fitted le` = yhat(leModel),
    `y fitted ek` = yhat(ekModel)
  ) %>% 
  select(Year, y, `y fitted kle`, `y fitted kl`, `y fitted le`, `y fitted ek`) %>% 
  gather(key = var, value = value, y, 
         `y fitted kle`, `y fitted kl`, `y fitted le`, `y fitted ek`) %>% 
  mutate(
    var = factor(var, levels = c("y", "y fitted kle", 
                                 "y fitted kl", "y fitted le", "y fitted ek"))
  ) %>% 
  ggplot(mapping = aes(x = Year, y = value, linetype = var, colour = var)) +
  geom_line() +
  geom_hline(yintercept = 1, colour = "gray50", size = 0.1) +
  labs(x = NULL, y = NULL, linetype = NULL, colour = NULL) +
  xy_theme()
```

We can also plot all of the residuals (which are calculated in log space).

```{r}
resids <- USData %>% 
  select(Year) %>% 
  mutate(
    r_y = 0,
    r_kle = residuals(kleModel) %>% as.numeric(),
    r_kl = residuals(klModel) %>% as.numeric(),
    r_le = residuals(leModel) %>% as.numeric(),
    r_ek = residuals(ekModel) %>% as.numeric()
  ) %>% 
  gather(key = var, value = value, r_y, r_kle, r_kl, r_le, r_ek) %>% 
  mutate(
    var = factor(var, levels = c("r_y", "r_kle", "r_kl", "r_le", "r_ek"))
  ) 
resids %>% 
  ggplot(mapping = aes(x = Year, y = value, linetype = var, colour = var)) +
  geom_line() + 
  labs(x = NULL, y = NULL, linetype = NULL, colour = NULL) +

The sum of squared residuals (in log space) can be used to compare the models.

```{r}
SSE <- resids %>% 
  mutate(
    r2 = value^2
  ) %>% 
  select(-value) %>% 
  spread(key = var, value = r2) %>% 
  summarise(
    sse_kle = sum(r_kle), 
    sse_kl = sum(r_kl),
    sse_le = sum(r_le),
    sse_ek = sum(r_ek)
  )
SSE
```

@@

Visually, the fit is quite good for all models. 
It is difficult to say that one model is vastly superior to any other model.
According to SSE, the $(kl)e$ model fits best, as would be expected.
